---
title: "Model Selection"
author: "AMF"
date: "2/14/2021"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Slash pine Model Selection
SMA regression was used to determine the relationship between log(Age) and log(DBH).
```{r sma, echo=FALSE, warning=FALSE}
setwd("C:/Users/Alicia/Documents/GitHub/FL_Carbon")

sitetree<-read.csv("sitetree_NAD83_envData.csv", header=T, sep=",", stringsAsFactors = F)



sitetree$SPCD<-as.factor(sitetree$SPCD)

# Calculate weighted averages and soil C:N for Soilgrids data -------------


is.na(sitetree$SOC0_5_NAD) <- sitetree$SOC0_5_NAD==-9999
is.na(sitetree$SOC5_15_NA) <- sitetree$SOC5_15_NA==-9999

is.na(sitetree$N0_5_NAD) <- sitetree$N0_5_NAD==-9999
is.na(sitetree$N5_15_NAD) <- sitetree$N5_15_NAD==-9999

is.na(sitetree$X30s_NAD) <- sitetree$X30s_NAD==-9999
is.na(sitetree$ai_et0_NAD) <- sitetree$ai_et0_NAD==-9999


sitetree$SOILGRIDS_C_AVG<-((sitetree$SOC0_5_NAD/10)*(1/3))+((sitetree$SOC5_15_NA/10)*(2/3))
                 

sitetree$SOILGRIDS_N_AVG<-((sitetree$N0_5_NAD/100)*(1/3))+((sitetree$N5_15_NAD/100)*(2/3))                 

sitetree$SOILGRIDS_CN<-sitetree$SOILGRIDS_C_AVG/sitetree$SOILGRIDS_N_AVG 


sitetree[which(is.finite(sitetree$SOILGRIDS_CN)==F),"SOILGRIDS_CN"]<-NA 

#scale CN from FIA data
sitetree$SOILGRIDS_CN_SCALE<-(sitetree$SOILGRIDS_CN*6.212)+24.634

sitetree$ai_et0_NAD<-sitetree$ai_et0_NAD*.0001

# Subsetting species ------------------------------------------------------

slashpine<-subset(sitetree, sitetree$SPCD=="111")
longleaf<-subset(sitetree, sitetree$SPCD=="121")
loblolly<-subset(sitetree, sitetree$SPCD=="131")
SlashMA <- smatr::sma(data=slashpine, log(slashpine$DIA)~log(slashpine$AGEDIA), method = "SMA")
slashresidualsma<-residuals(SlashMA)
summary(SlashMA)
```
## SMA regression line and residuals
```{r sma plot, echo=FALSE}
plot(SlashMA)
```
```{r resid plot, echo=FALSE}
slashresidualsma<-residuals(SlashMA)
plot(slashresidualsma)
```

Int.model3 was selected based upon the following criteria: lowest predicted error sum of squares (PRESS), highest Adjusted R-squared, and lowest BIC score.
```{r model criteria, echo=FALSE}

modelnames <- c("full.model", "add.model", "notemp.model", "int.model2", "int.model3", "int.model4")

PRESS <- c(117.2473, 121.8596, 122.7548, 121.7957, 116.8553, 117.4792)

Adj.r <- c(0.06016, 0.06254, 0.05519, 0.06339, 0.06364, 0.05821)

BIC <- c(-183.0376, -174.4199, -167.7221, -169.7606, -184.5131, -178.4704)

df<-cbind(modelnames, PRESS, Adj.r, BIC)

df<-data.frame(df)

colnames(df)<-c("Model","PRESS","Adj R-squared", "BIC")

print(df)
```

This is the selected model and its coefficients.
```{r final model, echo=FALSE}
int.model3<-lm(slashresidualsma~ ai_et0_NAD*SOILGRIDS_CN_SCALE + slashpine$X30s_NAD, data=slashpine)
summary(int.model3)
```

## Model Validation
Above-ground biomass (ABG) of 1 acre slash pine stands was simulated using individual-based growth and mortality model and compared to FIA remeasurement plots.
```{r slash simu, echo=FALSE}
rm(list=ls())

setwd("C:/Users/Alicia/Documents/GitHub/FL_Carbon/Slash Remeasurement")
envdata<-read.csv("SlashEnvData1.csv", header=T, sep=",") 
load("SlashAgeTotals.rdata")
load("SlashAgeTotalsEnd.rdata")
load("SlashDIATotalsEnd.rdata")
load("SlashDIATotals.rdata")
load("SlashPlotStart.rdata")
load("SlashPlotEnd.rdata")

mylist<-list()

simu.diameters<-list()

a<-c(1:18,20:21, 23:32, 34:56)

for (s in a){
  
  # set stand age and density
  plot_density<-diameter.totals[[s]]
  observed.a<-envdata[s,6]
  ages<-age.totals[[s]]
  temp<-envdata[s,7]
  CN<-envdata[s,16]
  aridity<-envdata[s,17]
  predict.tasb<-matrix(nrow = 10, ncol = 1,0)
  predict.d<-matrix(nrow = 10, ncol = 1,0)
  
  for (o in 1:10){
    age<-matrix(0, nrow=observed.a, ncol=length(plot_density)) # initialize the age matrix
    Diameter<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the diameter matrix
    TASB<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the total above-stump biomass matrix
    
    # initialize the diameter for the first year
    Diameter[1,]<-plot_density
    age[1,]<-ages
    
    for (j in 1:length(plot_density)){ # specify tree per hectare
      
      for (i in 2:observed.a){ # specify how long to run the simulation (years)
        
        age[i,j]<-age[i-1,j]+1
        
        growth<-(10^(2.103102 - 1.374381*aridity - 0.009940*CN  - 0.026373*temp +              0.010618*CN*aridity))*(0.5783973*age[i,j]^(-0.4216027))
        
        # define the mortality rate here
        # initialize as a numeric with only 1 possible value
        M <- numeric(length = 1)
        
        # Mortality based on diameter class
        if (Diameter[i,j]>=0) { M<- rbinom(1,1,(0.06504789*exp(-0.05577336*Diameter[i,j]*2.54)))}
        # if (Diameter[i,j]>0) { M<- rbinom(1,1,(exp(0.7800557-0.08071*Diameter[i,j]*2.54)/8))}
        # if (Diameter[i,j]<=3.94){ M<- rbinom(1,1,(.4/8))}
        # else if (Diameter[i,j]>3.94 & Diameter[i,j]<=7.97){M<-rbinom(1,1,(.2/8))}
        # else if (Diameter[i,j]>7.97 & Diameter[i,j]<=11.81){M<-rbinom(1,1,(.15/8))}
        # else if (Diameter[i,j]>11.81 & Diameter[i,j]<=15.75){M<-rbinom(1,1,(.1/8))}
        # else if (Diameter[i,j]>15.75){M<-rbinom(1,1,(0.01/8))}
        
        # Calculate the diameter for jth tree for the ith observed year
        Diameter[i,j]<-Diameter[i-1,j] + growth - M*(Diameter[i-1,j]+growth)
        
        #use diameter and age to calculate total aboveground biomass of the jth tree in the ith year
        TASB [i,j]<-(0.041281*((Diameter[i,j]*2.54)^2.722214))
        
        # If the tree dies, plant a new tree (age = 0)
        if (M==1){age[i,j]<-0}
      }
    }
    # save average modeled diameter
    predict.d[o,1]<-mean(Diameter[observed.a,])
    predict.tasb[o,1]<-sum(TASB[observed.a, ])*.5*(1/4047)
  }
  
  observed.d<-mean(diameter.totals.end[[s]])
  observed.tasb<-sum(plot_data_end[[s]]$TASB)*.5*(1/4047)
  modeled.d<-mean(predict.d)
  modeled.tasb<-mean(predict.tasb)
  sd.tasb<-sd(predict.tasb)
  sd.dbh<-sd(predict.d)
  # hist(Diameter[observed.a,], main=paste("End simulation", s), xlab = "Diameter (in)")
  
  # set up dataframe to store simulated data
  df<-cbind(observed.d, modeled.d, observed.a, length(plot_density), envdata[s,12], envdata[s,17], observed.tasb, modeled.tasb, 
            sd.tasb, sd.dbh)
  
  df<-data.frame(df)
  
  colnames(df)<-c("Observed_Diameter","Modeled_Diameter","Age", "Tree_Density", "Temperature",  "Aridity", "Observed_Biomass", 
                  "Modeled_Biomass", "sd.tasb", "sd.dbh")
  
  # data will be saved as list 1
  mylist[[s]] <- df
  simu.diameters[[s]]<-Diameter[observed.a,]
  
}

final_list <- do.call(rbind.data.frame, mylist)
row.names(final_list) <- c(a)
sdev<-as.vector(final_list$sd.tasb)
sdev.dbh<-as.vector(final_list$sd.dbh)

plot(data = final_list, Observed_Biomass~Modeled_Biomass, xlim = c(0,15), ylim = c(0,15), col = "#75BFBF", xlab="Modeled", ylab="Observed", main ="AGB (kgC/m^2)",
     col.axis="#027368", pch=16, type="p") 
abline(0,1, col="#048ABF")
# text(Observed_Biomass~Modeled_Biomass, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Biomass-sdev, final_list$Observed_Biomass, final_list$Modeled_Biomass+sdev, final_list$Observed_Biomass, length=0.05, angle=90, code=3)

plot(data = final_list, Observed_Diameter~Modeled_Diameter, xlim=c(0,20), ylim=c(0,20), xlab="Modeled DBH (in)", ylab="Observed DBH (in)",
      col.axis="#027368", col="#75BFBF", pch=16, type="p")
abline(0,1, col="#048ABF")
# text(Observed_Diameter~Modeled_Diameter, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Diameter-sdev.dbh, final_list$Observed_Diameter, final_list$Modeled_Diameter+sdev.dbh, final_list$Observed_Diameter, length=0.05, angle=90, code=3)
```

## Longleaf pine Model Selection
```{r, echo=FALSE}
setwd("C:/Users/Alicia/Documents/GitHub/FL_Carbon")

sitetree<-read.csv("sitetree_NAD83_envData.csv", header=T, sep=",", stringsAsFactors = F)



sitetree$SPCD<-as.factor(sitetree$SPCD)

# Calculate weighted averages and soil C:N for Soilgrids data -------------


is.na(sitetree$SOC0_5_NAD) <- sitetree$SOC0_5_NAD==-9999
is.na(sitetree$SOC5_15_NA) <- sitetree$SOC5_15_NA==-9999

is.na(sitetree$N0_5_NAD) <- sitetree$N0_5_NAD==-9999
is.na(sitetree$N5_15_NAD) <- sitetree$N5_15_NAD==-9999

is.na(sitetree$X30s_NAD) <- sitetree$X30s_NAD==-9999
is.na(sitetree$ai_et0_NAD) <- sitetree$ai_et0_NAD==-9999


sitetree$SOILGRIDS_C_AVG<-((sitetree$SOC0_5_NAD/10)*(1/3))+((sitetree$SOC5_15_NA/10)*(2/3))
                 

sitetree$SOILGRIDS_N_AVG<-((sitetree$N0_5_NAD/100)*(1/3))+((sitetree$N5_15_NAD/100)*(2/3))                 

sitetree$SOILGRIDS_CN<-sitetree$SOILGRIDS_C_AVG/sitetree$SOILGRIDS_N_AVG 


sitetree[which(is.finite(sitetree$SOILGRIDS_CN)==F),"SOILGRIDS_CN"]<-NA 

#scale CN from FIA data
sitetree$SOILGRIDS_CN_SCALE<-(sitetree$SOILGRIDS_CN*6.212)+24.634

sitetree$ai_et0_NAD<-sitetree$ai_et0_NAD*.0001

# Subsetting species ------------------------------------------------------

slashpine<-subset(sitetree, sitetree$SPCD=="111")
longleaf<-subset(sitetree, sitetree$SPCD=="121")
loblolly<-subset(sitetree, sitetree$SPCD=="131")

LongleafSMA <- smatr::sma(data=longleaf, log(DIA) ~ log(AGEDIA), method = "SMA")
summary(LongleafSMA)
```
## SMA Model and Residuals
```{r, echo=FALSE}
plot(LongleafSMA)
```
```{r, echo=FALSE}
resid.LLSMA<- residuals(LongleafSMA)
plot(resid.LLSMA)
```

Int.model3 was selected based upon the following criteria: lowest predicted error sum of squares (PRESS), highest Adjusted R-squared, and lowest BIC score.
```{r, echo=FALSE}
modelnames <- c("full.model", "add.model", "notemp.model", "int.model2", "int.model3", "int.model4", "int.model5")

PRESS <- c(35.60895, 36.59297, 37.60718, 35.76112,34.99162, 35.77264, 36.35493)

Adj.r <- c(0.1145, 0.07421, 0.0997, 0.1213, 0.1306, 0.1106, 0.09551)

BIC <- c(-124.0053, -117.941, -102.5656, -130.626, -132.0229, -120.7723, -113.8396)

df<-cbind(modelnames, PRESS, Adj.r, BIC)

df<-data.frame(df)

colnames(df)<-c("Model","PRESS","Adj R-squared", "BIC")

print(df)
```

This is the selected model and its coefficients.
```{r, echo=FALSE}
int.model3<-lm(resid.LLSMA~ X30s_NAD + ai_et0_NAD * SOILGRIDS_CN_SCALE, data=longleaf)
summary(int.model3)
```

## Model Diagnostics
```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(int.model3)
```

## Model Validation
Above-ground biomass (ABG) of 1 acre Longleaf pine stands was simulated using individual-based growth and mortality model and compared to FIA remeasurement plots. With probability of mortality by size class.
```{r, echo=FALSE, warning=FALSE}
rm(list=ls())

setwd("C:/Users/Alicia/Documents/GitHub/FL_Carbon/Longleaf Remeasurment")
envdata<-read.csv("LongleafEnvData1.csv", header=T, sep=",") 
load("longleafAgeTotals.rdata")
load("longleafAgeTotalsEnd.rdata")
load("longleafDIATotalsEnd.rdata")
load("longleafDIATotals.rdata")
load("longleafPlotStart.rdata")
load("longleafPlotEnd.rdata")
# load("initialpar300.Rdata")

par(mfrow=c(2,5))

mylist<-list()


# Plot 170 ----------------------------------------------

simu.diameters<-list()

a<-c(1, 3, 4, 6, 8, 9, 11:21, 23)
calibration<-c(21, 19, 16,  3, 23, 14,  9,  1, 20, 11)
a<-setdiff(a, calibration)

for (s in a){
  
  # set stand age and density
  plot_density<-diameter.totals[[s]]
  observed.a<-envdata[s,6]
  ages<-age.totals[[s]]
  CN <- envdata[s,16]
  aridity <- envdata[s,17]
  temp <- envdata[s,12]
  predict.tasb<-matrix(nrow = 300, ncol = 1,0)
  
  predict.d<-matrix(nrow = 300, ncol = 1,0)
  
  
  for (o in 1:300){
    age<-matrix(0, nrow=observed.a, ncol=length(plot_density)) # initialize the age matrix
    Diameter<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the diameter matrix
    TASB<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the total above-stump biomass matrix
    
    # initialize the diameter for the first year
    Diameter[1,]<-plot_density
    age[1,]<-ages
    
    for (j in 1:length(plot_density)){ # specify tree per hectare
      
      for (i in 2:observed.a){ # specify how long to run the simulation (years)
        
        age[i,j]<-age[i-1,j]+1
        
        growth<-(10^(4.085367 - 0.016985*CN + 0.020275*CN*aridity - 2.596493*aridity - 0.078576*temp))*(0.5718692*age[i,j]^(-0.4281308))
        
        # define the mortality rate here
        # initialize as a numeric with only 1 possible value
        M <- numeric(length = 1)
        
        # # Mortality based on diameter class
        if (Diameter[i,j]>0) {M<- rbinom(1,1,(2.109e-02 +(-1.048e-03 *Diameter[i,j]*2.54)+ 1.324e-05*((Diameter[i,j]*2.54)^2)))
        }

        
        # else {M=0}
        # if (Diameter[i,j]<=3.94){ M<- rbinom(1,1,(.162/8))}
        # else if (Diameter[i,j]>3.94 & Diameter[i,j]<=7.97){M<-rbinom(1,1,(.026/8))}
        # else if (Diameter[i,j]>7.97 & Diameter[i,j]<=11.81){M<-rbinom(1,1,(.006/8))}
        # else if (Diameter[i,j]>11.81 & Diameter[i,j]<=15.75){M<-rbinom(1,1,(.013/8))}
        # else if (Diameter[i,j]>15.75 & Diameter[i,j]<=19.69){M<-rbinom(1,1,(.024/8))}
        # else if (Diameter[i,j]>19.69 & Diameter[i,j]<=23.62){M<-rbinom(1,1,(.047/8))}
        # else if (Diameter[i,j]>23.62 & Diameter[i,j]<=27.56){M<-rbinom(1,1,(.060/8))}
        # else if (Diameter[i,j]>27.56){M<-rbinom(1,1,(0.129/8))}
        
        
        # Calculate the diameter for jth tree for the ith observed year
        Diameter[i,j]<-Diameter[i-1,j] + growth - M*(Diameter[i-1,j]+growth)
        
        #use diameter and age to calculate total aboveground biomass of the jth tree in the ith year
        TASB [i,j]<-(0.0725*((Diameter[i,j]*2.54)^2.5074))+(0.0016*((Diameter[i,j]*2.54)^3.0786))+(0.0214*((Diameter[i,j]*2.54)^2.0051))
        
        # If the tree dies, plant a new tree (age = 0)
        if (M==1){age[i,j]<-0}
      }
    }
    # save average modeled diameter
    predict.d[o,1]<-mean(Diameter[observed.a,])
    predict.tasb[o,1]<-sum(TASB[observed.a, ])*.5*(1/4047)
  
  }
  # mod.tasb[[s]]<-predict.tasb
  observed.d<-mean(diameter.totals.end[[s]])
  observed.tasb<-sum(plot_data_end[[s]]$TASB)*.5*(1/4047)
  modeled.d<-mean(predict.d)
  modeled.tasb<-mean(predict.tasb)
  # sd.tasb<-mad(predict.tasb, center = median(predict.tasb), constant = 1.4826, na.rm = FALSE,
      # low = FALSE, high = FALSE)
  sd.tasb<-sd(predict.tasb)
  sd.dbh<-sd(predict.d)
  # hist(Diameter[observed.a,], main=paste("End simulation", s), xlab = "Diameter (in)")
  
  # set up dataframe to store simulated data
  df<-cbind(observed.d, modeled.d, observed.a, length(plot_density), envdata[s,12], envdata[s,17], observed.tasb, modeled.tasb, 
            sd.tasb, sd.dbh)
  
  df<-data.frame(df)
  
  colnames(df)<-c("Observed_Diameter","Modeled_Diameter","Age", "Tree_Density", "Temperature",  "Aridity", "Observed_Biomass", 
                  "Modeled_Biomass", "sd.tasb", "sd.dbh")
  
  # data will be saved as list 1
  mylist[[s]] <- df
  simu.diameters[[s]]<-Diameter[observed.a,]
  # mod.tasb[[s]]<-predict.tasb[,1]
}

# par(mfrow=c(1,3))

# hist(diameter.totals[[5]])
# hist(diameter.totals.end[[5]])
# hist(simu.diameters[[5]])

# 

# outliers<-vector()
# for (m in a){
# outliers  <-boxplot(mod.tasb[[6]])$out
# }
# 
# which(mod.tasb[[6]] %in% outliers)
# mad(x, center = median(x), constant = 1.4826, na.rm = FALSE,
#     low = FALSE, high = FALSE)

final_list <- do.call(rbind.data.frame, mylist)
row.names(final_list) <- c(a)
sdev<-as.vector(final_list$sd.tasb)
sdev.dbh<-as.vector(final_list$sd.dbh)

par(mfrow=c(1,1))


plot(data = final_list, Observed_Biomass~Modeled_Biomass,col = "#75BFBF", xlab="Modeled", ylab="Observed", main ="AGB before data assimilation (kgC/m^2)",
     col.axis="#027368", pch=16, type="p")
abline(0,1, col="#048ABF")
text(Observed_Biomass~Modeled_Biomass, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Biomass-sdev, final_list$Observed_Biomass, final_list$Modeled_Biomass+sdev, final_list$Observed_Biomass, length=0.05, angle=90, code=3)

plot(data = final_list, Observed_Diameter~Modeled_Diameter, xlim=c(0,20), ylim=c(0,20), xlab="Modeled DBH (in)", ylab="Observed DBH (in)", main ="Before data assimilation",
      col.axis="#027368", col="#75BFBF", pch=16, type="p")
abline(0,1, col="#048ABF")
# text(Observed_Diameter~Modeled_Diameter, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Diameter-sdev.dbh, final_list$Observed_Diameter, final_list$Modeled_Diameter+sdev.dbh, final_list$Observed_Diameter, length=0.05, angle=90, code=3)
```
## Longleaf after data assimilation
Running 300 random parameter sets (after discarding burn-in) per site. Reporting site means with site sd

```{r, echo=FALSE}
rm(list=ls())

setwd("C:/Users/Alicia/Documents/GitHub/FL_Carbon/Longleaf Remeasurment")
envdata<-read.csv("LongleafEnvData1.csv", header=T, sep=",") 
load("longleafAgeTotals.rdata")
load("longleafAgeTotalsEnd.rdata")
load("longleafDIATotalsEnd.rdata")
load("longleafDIATotals.rdata")
load("longleafPlotStart.rdata")
load("longleafPlotEnd.rdata")
load("sampleparameters2.20.21.rdata")
# load("LLMortTest300par.rdata")

par(mfrow=c(2,5))

mylist<-list()
mod.tasb<-list()
simu.diameters<-list()

a<-c(1, 3, 4, 6, 8, 9, 11:21, 23)
calibration<-c(21, 19, 16,  3, 23, 14,  9,  1, 20, 11)
a<-setdiff(a, calibration)

# a<-c(1, 3:23)

for (s in a){
  
  # set stand age and density
  plot_density<-diameter.totals[[s]]
  observed.a<-envdata[s,6]
  ages<-age.totals[[s]]
  CN <- envdata[s,16]
  aridity <- envdata[s,17]
  temp <- envdata[s,12]
  predict.tasb<-matrix(nrow = 300, ncol = 1,0)
  
  predict.d<-matrix(nrow = 300, ncol = 1,0)
  
  
  for (o in 1:300){
    age<-matrix(0, nrow=observed.a, ncol=length(plot_density)) # initialize the age matrix
    Diameter<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the diameter matrix
    TASB<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the total above-stump biomass matrix
    
    # initialize the diameter for the first year
    Diameter[1,]<-plot_density
    age[1,]<-ages
    
    for (j in 1:length(plot_density)){ # specify tree per hectare
      
      for (i in 2:observed.a){ # specify how long to run the simulation (years)
        
        age[i,j]<-age[i-1,j]+1
        
        growth<-(10^(sample.parameters[1,o] - sample.parameters[2,o]*CN + sample.parameters[3,o]*CN*aridity - sample.parameters[4,o]*aridity- sample.parameters[5,o]*temp))*(sample.parameters[6,o]*age[i,j]^(sample.parameters[7,o]))
        
        # define the mortality rate here
        # initialize as a numeric with only 1 possible value
        M <- numeric(length = 1)
        
        # # Mortality based on diameter class
        if (Diameter[i,j]>0) {
          M<- rbinom(1,1,(sample.parameters[8,o] +(sample.parameters[9,o] *Diameter[i,j]*2.54)+sample.parameters[10,o]*((Diameter[i,j]*2.54)^2)))

        }
        # if (Diameter[i,j]<=3.94){ M<- rbinom(1,1,(.162/8))}
        # else if (Diameter[i,j]>3.94 & Diameter[i,j]<=7.97){M<-rbinom(1,1,(.026/8))}
        # else if (Diameter[i,j]>7.97 & Diameter[i,j]<=11.81){M<-rbinom(1,1,(.006/8))}
        # else if (Diameter[i,j]>11.81 & Diameter[i,j]<=15.75){M<-rbinom(1,1,(.013/8))}
        # else if (Diameter[i,j]>15.75 & Diameter[i,j]<=19.69){M<-rbinom(1,1,(.024/8))}
        # else if (Diameter[i,j]>19.69 & Diameter[i,j]<=23.62){M<-rbinom(1,1,(.047/8))}
        # else if (Diameter[i,j]>23.62 & Diameter[i,j]<=27.56){M<-rbinom(1,1,(.060/8))}
        # else if (Diameter[i,j]>27.56){M<-rbinom(1,1,(0.129/8))}
        
        
        # Calculate the diameter for jth tree for the ith observed year
        Diameter[i,j]<-Diameter[i-1,j] + growth - M*(Diameter[i-1,j]+growth)
        
        #use diameter and age to calculate total aboveground biomass of the jth tree in the ith year
        TASB [i,j]<-(0.0725*((Diameter[i,j]*2.54)^2.5074))+(0.0016*((Diameter[i,j]*2.54)^3.0786))+(0.0214*((Diameter[i,j]*2.54)^2.0051))
        
        # If the tree dies, plant a new tree (age = 0)
        if (M==1){age[i,j]<-0}
      }
    }
    # save average modeled diameter
    predict.d[o,1]<-mean(Diameter[observed.a,])
    predict.tasb[o,1]<-sum(TASB[observed.a, ])*.5*(1/4047)
  
  }
  # mod.tasb[[s]]<-predict.tasb
  observed.d<-mean(diameter.totals.end[[s]])
  observed.tasb<-sum(plot_data_end[[s]]$TASB)*.5*(1/4047)
  modeled.d<-mean(predict.d)
  modeled.tasb<-mean(predict.tasb)
  # sd.tasb<-mad(predict.tasb, center = median(predict.tasb), constant = 1.4826, na.rm = FALSE,
      # low = FALSE, high = FALSE)
  sd.tasb<-sd(predict.tasb)
  sd.dbh<-sd(predict.d)
  # hist(Diameter[observed.a,], main=paste("End simulation", s), xlab = "Diameter (in)")
  
  # set up dataframe to store simulated data
  df<-cbind(observed.d, modeled.d, observed.a, length(plot_density), envdata[s,12], envdata[s,17], observed.tasb, modeled.tasb, 
            sd.tasb, sd.dbh)
  
  df<-data.frame(df)
  
  colnames(df)<-c("Observed_Diameter","Modeled_Diameter","Age", "Tree_Density", "Temperature",  "Aridity", "Observed_Biomass", 
                  "Modeled_Biomass", "sd.tasb", "sd.dbh")
  
  # data will be saved as list 1
  mylist[[s]] <- df
  simu.diameters[[s]]<-Diameter[observed.a,]
  # mod.tasb[[s]]<-predict.tasb[,1]
}

# par(mfrow=c(1,3))

# hist(diameter.totals[[5]])
# hist(diameter.totals.end[[5]])
# hist(simu.diameters[[5]])

# 

# outliers<-vector()
# for (m in a){
# outliers  <-boxplot(mod.tasb[[6]])$out
# }
# 
# which(mod.tasb[[6]] %in% outliers)
# mad(x, center = median(x), constant = 1.4826, na.rm = FALSE,
#     low = FALSE, high = FALSE)

final_list <- do.call(rbind.data.frame, mylist)
row.names(final_list) <- c(a)
sdev<-as.vector(final_list$sd.tasb)
sdev.dbh<-as.vector(final_list$sd.dbh)

par(mfrow=c(1,1))


plot(data = final_list, Observed_Biomass~Modeled_Biomass,col = "#75BFBF", xlab="Modeled", ylab="Observed", main ="AGB after data assimilation (kgC/m^2)",
     col.axis="#027368", pch=16, type="p")
abline(0,1, col="#048ABF")
text(Observed_Biomass~Modeled_Biomass, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Biomass-sdev, final_list$Observed_Biomass, final_list$Modeled_Biomass+sdev, final_list$Observed_Biomass, length=0.05, angle=90, code=3)

plot(data = final_list, Observed_Diameter~Modeled_Diameter, xlim=c(0,20), ylim=c(0,20), xlab="Modeled DBH (in)", ylab="Observed DBH (in)", main ="After data assimilation",
      col.axis="#027368", col="#75BFBF", pch=16, type="p")
abline(0,1, col="#048ABF")
# text(Observed_Diameter~Modeled_Diameter, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Diameter-sdev.dbh, final_list$Observed_Diameter, final_list$Modeled_Diameter+sdev.dbh, final_list$Observed_Diameter, length=0.05, angle=90, code=3)

```
## Model Validation
Above-ground biomass (ABG) of 1 acre Longleaf pine stands was simulated using individual-based growth and mortality model and compared to FIA remeasurement plots. With continuous probability of mortality by size class.
```{r, echo=FALSE}
rm(list=ls())

setwd("C:/Users/Alicia/Documents/GitHub/FL_Carbon/Longleaf Remeasurment")
envdata<-read.csv("LongleafEnvData1.csv", header=T, sep=",") 
load("longleafAgeTotals.rdata")
load("longleafAgeTotalsEnd.rdata")
load("longleafDIATotalsEnd.rdata")
load("longleafDIATotals.rdata")
load("longleafPlotStart.rdata")
load("longleafPlotEnd.rdata")
# load("initialpar300.Rdata")

par(mfrow=c(2,5))

mylist<-list()


# Plot 170 ----------------------------------------------

simu.diameters<-list()

a<-c(1, 3, 4, 6, 8, 9, 11:21, 23)
calibration<-c(21, 19, 16,  3, 23, 14,  9,  1, 20, 11)
a<-setdiff(a, calibration)

for (s in a){
  
  # set stand age and density
  plot_density<-diameter.totals[[s]]
  observed.a<-envdata[s,6]
  ages<-age.totals[[s]]
  CN <- envdata[s,16]
  aridity <- envdata[s,17]
  temp <- envdata[s,12]
  predict.tasb<-matrix(nrow = 300, ncol = 1,0)
  
  predict.d<-matrix(nrow = 300, ncol = 1,0)
  
  
  for (o in 1:300){
    age<-matrix(0, nrow=observed.a, ncol=length(plot_density)) # initialize the age matrix
    Diameter<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the diameter matrix
    TASB<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the total above-stump biomass matrix
    
    # initialize the diameter for the first year
    Diameter[1,]<-plot_density
    age[1,]<-ages
    
    for (j in 1:length(plot_density)){ # specify tree per hectare
      
      for (i in 2:observed.a){ # specify how long to run the simulation (years)
        
        age[i,j]<-age[i-1,j]+1
        
        growth<-(10^(4.085367 - 0.016985*CN + 0.020275*CN*aridity - 2.596493*aridity - 0.078576*temp))*(0.5718692*age[i,j]^(-0.4281308))
        
        # define the mortality rate here
        # initialize as a numeric with only 1 possible value
        M <- numeric(length = 1)
        
        # # Mortality based on diameter class
        if (Diameter[i,j]>=0) {M<- rbinom(1,1,(2.109e-02 +(-1.048e-03 *Diameter[i,j]*2.54)+ 1.324e-05*((Diameter[i,j]*2.54)^2)))
        }

        
        # else {M=0}
        # if (Diameter[i,j]<=3.94){ M<- rbinom(1,1,(.162/8))}
        # else if (Diameter[i,j]>3.94 & Diameter[i,j]<=7.97){M<-rbinom(1,1,(.026/8))}
        # else if (Diameter[i,j]>7.97 & Diameter[i,j]<=11.81){M<-rbinom(1,1,(.006/8))}
        # else if (Diameter[i,j]>11.81 & Diameter[i,j]<=15.75){M<-rbinom(1,1,(.013/8))}
        # else if (Diameter[i,j]>15.75 & Diameter[i,j]<=19.69){M<-rbinom(1,1,(.024/8))}
        # else if (Diameter[i,j]>19.69 & Diameter[i,j]<=23.62){M<-rbinom(1,1,(.047/8))}
        # else if (Diameter[i,j]>23.62 & Diameter[i,j]<=27.56){M<-rbinom(1,1,(.060/8))}
        # else if (Diameter[i,j]>27.56){M<-rbinom(1,1,(0.129/8))}
        
        
        # Calculate the diameter for jth tree for the ith observed year
        Diameter[i,j]<-Diameter[i-1,j] + growth - M*(Diameter[i-1,j]+growth)
        
        #use diameter and age to calculate total aboveground biomass of the jth tree in the ith year
        TASB [i,j]<-(0.0725*((Diameter[i,j]*2.54)^2.5074))+(0.0016*((Diameter[i,j]*2.54)^3.0786))+(0.0214*((Diameter[i,j]*2.54)^2.0051))
        
        # If the tree dies, plant a new tree (age = 0)
        if (M==1){age[i,j]<-0}
      }
    }
    # save average modeled diameter
    predict.d[o,1]<-mean(Diameter[observed.a,])
    predict.tasb[o,1]<-sum(TASB[observed.a, ])*.5*(1/4047)
  
  }
  # mod.tasb[[s]]<-predict.tasb
  observed.d<-mean(diameter.totals.end[[s]])
  observed.tasb<-sum(plot_data_end[[s]]$TASB)*.5*(1/4047)
  modeled.d<-mean(predict.d)
  modeled.tasb<-mean(predict.tasb)
  # sd.tasb<-mad(predict.tasb, center = median(predict.tasb), constant = 1.4826, na.rm = FALSE,
      # low = FALSE, high = FALSE)
  sd.tasb<-sd(predict.tasb)
  sd.dbh<-sd(predict.d)
  # hist(Diameter[observed.a,], main=paste("End simulation", s), xlab = "Diameter (in)")
  
  # set up dataframe to store simulated data
  df<-cbind(observed.d, modeled.d, observed.a, length(plot_density), envdata[s,12], envdata[s,17], observed.tasb, modeled.tasb, 
            sd.tasb, sd.dbh)
  
  df<-data.frame(df)
  
  colnames(df)<-c("Observed_Diameter","Modeled_Diameter","Age", "Tree_Density", "Temperature",  "Aridity", "Observed_Biomass", 
                  "Modeled_Biomass", "sd.tasb", "sd.dbh")
  
  # data will be saved as list 1
  mylist[[s]] <- df
  simu.diameters[[s]]<-Diameter[observed.a,]
  # mod.tasb[[s]]<-predict.tasb[,1]
}

# par(mfrow=c(1,3))

# hist(diameter.totals[[5]])
# hist(diameter.totals.end[[5]])
# hist(simu.diameters[[5]])

# 

# outliers<-vector()
# for (m in a){
# outliers  <-boxplot(mod.tasb[[6]])$out
# }
# 
# which(mod.tasb[[6]] %in% outliers)
# mad(x, center = median(x), constant = 1.4826, na.rm = FALSE,
#     low = FALSE, high = FALSE)

final_list <- do.call(rbind.data.frame, mylist)
row.names(final_list) <- c(a)
sdev<-as.vector(final_list$sd.tasb)
sdev.dbh<-as.vector(final_list$sd.dbh)

par(mfrow=c(1,1))


plot(data = final_list, Observed_Biomass~Modeled_Biomass,col = "#75BFBF", xlab="Modeled", ylab="Observed", main ="AGB before data assimilation (kgC/m^2)",
     col.axis="#027368", pch=16, type="p")
abline(0,1, col="#048ABF")
text(Observed_Biomass~Modeled_Biomass, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Biomass-sdev, final_list$Observed_Biomass, final_list$Modeled_Biomass+sdev, final_list$Observed_Biomass, length=0.05, angle=90, code=3)

plot(data = final_list, Observed_Diameter~Modeled_Diameter, xlim=c(0,20), ylim=c(0,20), xlab="Modeled DBH (in)", ylab="Observed DBH (in)", main ="Before data assimilation",
      col.axis="#027368", col="#75BFBF", pch=16, type="p")
abline(0,1, col="#048ABF")
# text(Observed_Diameter~Modeled_Diameter, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Diameter-sdev.dbh, final_list$Observed_Diameter, final_list$Modeled_Diameter+sdev.dbh, final_list$Observed_Diameter, length=0.05, angle=90, code=3)
```

## Longleaf after data assimilation
Running 300 random parameter sets (after discarding burn-in) per site using >=mortality function and sd denominator 4. Reporting site means with site sd

```{r, echo=FALSE}
rm(list=ls())

setwd("C:/Users/Alicia/Documents/GitHub/FL_Carbon/Longleaf Remeasurment")
envdata<-read.csv("LongleafEnvData1.csv", header=T, sep=",") 
load("longleafAgeTotals.rdata")
load("longleafAgeTotalsEnd.rdata")
load("longleafDIATotalsEnd.rdata")
load("longleafDIATotals.rdata")
load("longleafPlotStart.rdata")
load("longleafPlotEnd.rdata")
# load("sampleparameters2.20.21.rdata")
load("LLMortTest300par.rdata")

par(mfrow=c(2,5))

mylist<-list()
mod.tasb<-list()
simu.diameters<-list()

a<-c(1, 3, 4, 6, 8, 9, 11:21, 23)
calibration<-c(21, 19, 16,  3, 23, 14,  9,  1, 20, 11)
a<-setdiff(a, calibration)

# a<-c(1, 3:23)

for (s in a){
  
  # set stand age and density
  plot_density<-diameter.totals[[s]]
  observed.a<-envdata[s,6]
  ages<-age.totals[[s]]
  CN <- envdata[s,16]
  aridity <- envdata[s,17]
  temp <- envdata[s,12]
  predict.tasb<-matrix(nrow = 300, ncol = 1,0)
  
  predict.d<-matrix(nrow = 300, ncol = 1,0)
  
  
  for (o in 1:300){
    age<-matrix(0, nrow=observed.a, ncol=length(plot_density)) # initialize the age matrix
    Diameter<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the diameter matrix
    TASB<-matrix(0, nrow=observed.a, length(plot_density)) # initialize the total above-stump biomass matrix
    
    # initialize the diameter for the first year
    Diameter[1,]<-plot_density
    age[1,]<-ages
    
    for (j in 1:length(plot_density)){ # specify tree per hectare
      
      for (i in 2:observed.a){ # specify how long to run the simulation (years)
        
        age[i,j]<-age[i-1,j]+1
        
        growth<-(10^(sample.parameters[1,o] - sample.parameters[2,o]*CN + sample.parameters[3,o]*CN*aridity - sample.parameters[4,o]*aridity- sample.parameters[5,o]*temp))*(sample.parameters[6,o]*age[i,j]^(sample.parameters[7,o]))
        
        # define the mortality rate here
        # initialize as a numeric with only 1 possible value
        M <- numeric(length = 1)
        
        # # Mortality based on diameter class
        if (Diameter[i,j]>=0) {
          M<- rbinom(1,1,(sample.parameters[8,o] +(sample.parameters[9,o] *Diameter[i,j]*2.54)+sample.parameters[10,o]*((Diameter[i,j]*2.54)^2)))

        }
        # if (Diameter[i,j]<=3.94){ M<- rbinom(1,1,(.162/8))}
        # else if (Diameter[i,j]>3.94 & Diameter[i,j]<=7.97){M<-rbinom(1,1,(.026/8))}
        # else if (Diameter[i,j]>7.97 & Diameter[i,j]<=11.81){M<-rbinom(1,1,(.006/8))}
        # else if (Diameter[i,j]>11.81 & Diameter[i,j]<=15.75){M<-rbinom(1,1,(.013/8))}
        # else if (Diameter[i,j]>15.75 & Diameter[i,j]<=19.69){M<-rbinom(1,1,(.024/8))}
        # else if (Diameter[i,j]>19.69 & Diameter[i,j]<=23.62){M<-rbinom(1,1,(.047/8))}
        # else if (Diameter[i,j]>23.62 & Diameter[i,j]<=27.56){M<-rbinom(1,1,(.060/8))}
        # else if (Diameter[i,j]>27.56){M<-rbinom(1,1,(0.129/8))}
        
        
        # Calculate the diameter for jth tree for the ith observed year
        Diameter[i,j]<-Diameter[i-1,j] + growth - M*(Diameter[i-1,j]+growth)
        
        #use diameter and age to calculate total aboveground biomass of the jth tree in the ith year
        TASB [i,j]<-(0.0725*((Diameter[i,j]*2.54)^2.5074))+(0.0016*((Diameter[i,j]*2.54)^3.0786))+(0.0214*((Diameter[i,j]*2.54)^2.0051))
        
        # If the tree dies, plant a new tree (age = 0)
        if (M==1){age[i,j]<-0}
      }
    }
    # save average modeled diameter
    predict.d[o,1]<-mean(Diameter[observed.a,])
    predict.tasb[o,1]<-sum(TASB[observed.a, ])*.5*(1/4047)
  
  }
  # mod.tasb[[s]]<-predict.tasb
  observed.d<-mean(diameter.totals.end[[s]])
  observed.tasb<-sum(plot_data_end[[s]]$TASB)*.5*(1/4047)
  modeled.d<-mean(predict.d)
  modeled.tasb<-mean(predict.tasb)
  # sd.tasb<-mad(predict.tasb, center = median(predict.tasb), constant = 1.4826, na.rm = FALSE,
      # low = FALSE, high = FALSE)
  sd.tasb<-sd(predict.tasb)
  sd.dbh<-sd(predict.d)
  # hist(Diameter[observed.a,], main=paste("End simulation", s), xlab = "Diameter (in)")
  
  # set up dataframe to store simulated data
  df<-cbind(observed.d, modeled.d, observed.a, length(plot_density), envdata[s,12], envdata[s,17], observed.tasb, modeled.tasb, 
            sd.tasb, sd.dbh)
  
  df<-data.frame(df)
  
  colnames(df)<-c("Observed_Diameter","Modeled_Diameter","Age", "Tree_Density", "Temperature",  "Aridity", "Observed_Biomass", 
                  "Modeled_Biomass", "sd.tasb", "sd.dbh")
  
  # data will be saved as list 1
  mylist[[s]] <- df
  simu.diameters[[s]]<-Diameter[observed.a,]
  # mod.tasb[[s]]<-predict.tasb[,1]
}

# par(mfrow=c(1,3))

# hist(diameter.totals[[5]])
# hist(diameter.totals.end[[5]])
# hist(simu.diameters[[5]])

# 

# outliers<-vector()
# for (m in a){
# outliers  <-boxplot(mod.tasb[[6]])$out
# }
# 
# which(mod.tasb[[6]] %in% outliers)
# mad(x, center = median(x), constant = 1.4826, na.rm = FALSE,
#     low = FALSE, high = FALSE)

final_list <- do.call(rbind.data.frame, mylist)
row.names(final_list) <- c(a)
sdev<-as.vector(final_list$sd.tasb)
sdev.dbh<-as.vector(final_list$sd.dbh)

par(mfrow=c(1,1))


plot(data = final_list, Observed_Biomass~Modeled_Biomass,col = "#75BFBF", xlab="Modeled", ylab="Observed", main ="AGB after data assimilation (kgC/m^2)",
     col.axis="#027368", pch=16, type="p")
abline(0,1, col="#048ABF")
text(Observed_Biomass~Modeled_Biomass, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Biomass-sdev, final_list$Observed_Biomass, final_list$Modeled_Biomass+sdev, final_list$Observed_Biomass, length=0.05, angle=90, code=3)

plot(data = final_list, Observed_Diameter~Modeled_Diameter, xlim=c(0,20), ylim=c(0,20), xlab="Modeled DBH (in)", ylab="Observed DBH (in)", main ="After data assimilation",
      col.axis="#027368", col="#75BFBF", pch=16, type="p")
abline(0,1, col="#048ABF")
# text(Observed_Diameter~Modeled_Diameter, labels=rownames(final_list),data=final_list, cex=0.9, font=2, pos=4)
arrows(final_list$Modeled_Diameter-sdev.dbh, final_list$Observed_Diameter, final_list$Modeled_Diameter+sdev.dbh, final_list$Observed_Diameter, length=0.05, angle=90, code=3)

```